---
title: "Maximum Approximate Bernstein/Beta Likelihood Estimation in R package 'mable'"
author: "Zhong Guan"
date: "02/08/2021"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Maximum Approximate Bernstein/Beta Likelihood Estimation in R package 'mable'}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
header-includes: \usepackage{bm}
link-citations: yes
bibliography:
- bernstein.bib
- por-model.bib
- chpt.bib
- wieand.bib
biblio-style: chicago
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


\setcounter{section}{0}
# Introduction

Any continuous density function $f$ on a known closed interval $[a,b]$ can be approximated by Bernstein polynomial $f_m(x; p)=(b-a)^{-1}\sum_{i=0}^m p_i\beta_{mi}[(x-a)/(b-a)]$, where $p_i\ge 0$, $\sum_{i=0}^m p_i=1$ and $\beta_{mi}(u)=(m+1){m\choose i}u^i(1-u)^{m-i}$, $i=0,1,\ldots,m$, is the beta density with shapes $(i+1, m-i+1)$. This provides a way to approximately model the unknwon underlying density with a mixture beta model with an appropriate \emph{model degree} $m$ and solve a nonparametric or semiparametric statistical problem ``parametrically'' using the maximum likelihood method. For instance, based on one-sample data, $x_1,\ldots,x_n$, one can estimate a nonparametric density $f$ parametrically by maximizing the approximate likelihood $\ell(p)=\sum_{j=0}^n\log f_m(x_j; p)$ with respect to $p$ [@Guan-jns-2015]. 

Since the Bernstein polynomial model of degree $m$ is nested in the model of degree $m+1$, the maximum likelihood is increasing in $m$. The increase is negligible when the model becomes overfitting. Therefore an optimal degree can be chosen as the change-point of the log likelihood ratios over a set of consecutive candidate model degrees. 

This approach works surprisingly well for even more complicated models and data. With an estimate $\hat p=(\hat p_0,\ldots,\hat p_m)$ of $p$ one can estimate the cumulative distribution function $F$ by
$\hat F(x)=F_m(x;\hat p)=\sum_{i=0}^m \hat p_i B_{mi}[(x-a)/(b-a)]$, where $B_{mi}(u)=\int_0^u\beta_{mj}(t)dt$, $i=0,\ldots,m$, is the beta distribution function with shapes $(i+1,m-i+1)$. This manual will illustrate the use of the R package \texttt{mable} for obtaining not only smooth estimates of density, cumulative distribution, and survival functions but also estimates of parameters such as regression coefficients. 

```{r echo = FALSE}
library(mable)
```

# One-sample Problems 
## Raw Data
Let $x_1,\ldots,x_n$ be a sample from a population with cumulative distribution function $F$ and density function $f$ on $[a,b]$. If $[a,b]$ is unknown we choose $[a,b]\supset [x_{(1)},x_{(n)}]$, where $x_{(1)}$ and $x_{(n)}$ are the minimum and maximum statistics, respectively. 

### Example: Vaal River Annual Flow Data
For the annual flow data of Vaal River at Standerton as given by Table 1.1 of @Linhart-and-Zucchini-model-selection-book give the flow in millions of cubic metres, 
```{r}
data(Vaal.Flow)
head(Vaal.Flow, 3)
```
we want to estimate the density and the distribution functions of annual flow
```{r results = "hide", warning = FALSE, message=FALSE}
vaal<-mable(Vaal.Flow$Flow, M = c(2,100), interval = c(0, 3000), IC = "all",
       controls = mable.ctrl(sig.level = 1e-8, maxit = 2000, eps = 1.0e-9))
```
Here we truncate the density by \texttt{interval=c(0, 3000)} and choose an optimal degree $m$ among the candidate degrees M[1]:M[2] using the method of change-point. The maximum number of iterations is \texttt{maxit} and the convergence criterion is \texttt{eps} for each $m$ of M[1]:M[2]. The search of an optimal degree stops when the $p$-value \texttt{pval} of change-point test falls below the specified significance level \texttt{sig.level} or the largest degree, M[2], has been reached. If the latter occurs a warning message shows up. In this case we should check the last value of \texttt{pval}. In the above example, we got warning message and the last \texttt{pval}, `r vaal$pval[vaal$M[2]-vaal$M[1]+1]`, which is small enough. The selected optimal degree is $m=$ `r vaal$m[1]`. One can also look at the Bayesian information criteria, \texttt{BIC}, and other information criteria, Akaike information criterion \texttt{AIC} and Hannan–Quinn information criterion \texttt{QHC}, at each candidate degree. These information criteria are not reliable due to the difficulty of determining the model dimension. The \texttt{plot} method for \texttt{mable} class object can visualize some of the results. 
```{r fig.align='center', fig.cap="Vaal River Annual Flow Data\\label{fig:vaal-river-data-plot}", fig.width=7, fig.height=7}
op <- par(mfrow = c(1,2))
layout(rbind(c(1, 2), c(3, 3)))
plot(vaal, which = "likelihood", cex = .5)
plot(vaal, which = "change-point", lgd.x = "topright")
hist(Vaal.Flow$Flow, prob = TRUE, xlim = c(0,3000), ylim =c(0,.0022), breaks =100*(0:30), 
    main = "Histogram and Densities of the Annual Flow of Vaal River",
    border = "dark grey",lwd = 1, xlab = "Flow", ylab = "Density", col ="light grey")
lines(density(x<-Vaal.Flow$Flow, bw = "nrd0", adjust = 1), lty = 2, col = 2,lwd = 2)
lines(y<-seq(0, 3000, length=100), dlnorm(y, mean(log(x)), sqrt(var(log(x)))), 
    lty = 4, col = 4, lwd = 2)
plot(vaal, which = "density", add = TRUE, lwd = 2)
legend("topright", lty = c(1, 4, 2), col = c(1, 4, 2), bty = "n",lwd = 2, 
c(expression(paste("MABLE: ",hat(f)[B])), expression(paste("Log-Normal: ",hat(f)[P])),
    expression(paste("KDE: ",hat(f)[K]))))
par(op)
```
In Figure \ref{fig:vaal-river-data-plot}, the unknown density $f$ is estimated by \texttt{MABLE}  $\hat f_B$ using optimal degrees $m=$ `r vaal$m` selected using the exponential change-point method, the parametric estimate using \texttt{Log-Normal} model and the kernel density estimate \texttt{KDE}: $\hat f_K$.

We can also look at the plots of AIC, BIC, and QHC, and likelihood ration(LR) of gamma change-point. 
```{r fig.align='center', fig.cap="AIC and BIC based on Vaal River Data\\label{fig:Vaal-River-data-AIC-BIC-plot}", fig.width=6, fig.height=4}
M <- vaal$M[1]:vaal$M[2]
aic <- vaal$ic$AIC
bic <- vaal$ic$BIC
qhc <- vaal$ic$QHC
vaal.gcp <- optim.gcp(vaal) # choose m by gamma change-point model
lr <- vaal.gcp$lr
plot(M, aic, cex = 0.7, col = 1, xlab = "m", ylab = "", ylim = c(ymin<-min(aic,bic,qhc,lr),
      ymax<-max(aic,bic,qhc,lr)), main = "AIC, BIC, QHC, and LR")
points(M, bic, pch = 2, cex = 0.7, col = 2)
points(M, qhc, pch = 3, cex = 0.7, col = 3)
points(M[-1], lr, pch = 4, cex = 0.7, col = 4)
segments(which.max(aic)+M[1]-1->m1, ymin, m1, max(aic), lty = 2)
segments(which.max(bic)+M[1]-1->m2, ymin, m2, max(bic), lty = 2, col = 2)
segments(which.max(qhc)+M[1]-1->m3, ymin, m3, max(qhc), lty = 2, col = 3)
segments(which.max(lr)+M[1]->m4, ymin, m4, max(lr), lty = 2, col = 4)
axis(1, c(m1,m2, m3, m4),  as.character(c(m1,m2,m3,m4)), col.axis = 4)
legend("topright", pch=c(1,2,3,4), c("AIC", "BIC", "QHC", "LR"), bty="n", col=c(1,2,3,4))
```
From Figure \ref{fig:Vaal-River-data-AIC-BIC-plot} we see that the gamma change-point method gives the same optimal degree $m=$ `r m4` as the exponential method; BIC and QHC give the same degree $m=$ `r m2`; the degree $m=$ `r m1` selected by AIC method is closer to the one selected by change-point methods. From this Figure we also see that unlike the LR plot the information criteria do not have clear peak points.

For any given degree $m$, one can fit the data by specifying \texttt{M=m} in \texttt{mable()} to obtain an estimate of $f$.

The \texttt{summary} method \texttt{summary.mable} prints and returns invisibly the summarized results.

```{r}
summary(vaal)
```
The mixing proportions, the coefficients of the Bernstein polynomials, $p$ can be obtained as either \texttt{p<-vaal\$p} 
```{r}
p<-vaal$p
p
```
or \texttt{summary(res)\$p}. 

## Grouped Data

With a grouped dataset, a frequency table of data from a continuous population, one can estimate the density from histogram using \texttt{mable.group()} with an optimal degree $m$ chosen from \texttt{M[1]:M[2]} or with a given degree $m$ using \texttt{M=m} [@Guan-2017-jns].

### Example: The Chicken Embryo Data
Consider the chicken embryo data contain the number of hatched eggs on each day during the 21 days of incubation period. The times of hatching ($n=43$) are treated as grouped by intervals with equal width of one day. The data were studied first by @Jassim-et-al-1996. @Kuurman-et-al-2003 and @Lin-and-He-2006-bka also analyzed the data using the minimum Hellinger distance estimator, in addition to other methods assuming some parametric mixture models including Weibull model. 

```{r}
data(chicken.embryo)
head(chicken.embryo, 2)
```
```{r results = "hide", warning = FALSE}
a <- 0
b <- 21
day <- chicken.embryo$day
nT <- chicken.embryo$nT
embryo<-mable.group(x = nT, breaks = a:b, M=c(2,100), interval = c(a, b), IC = "aic",
    controls = mable.ctrl(sig.level = 1e-6,  maxit = 2000, eps = 1.0e-7))
```

```{r fig.align='center', fig.cap="Chicken Embryo Data\\label{fig:chicken-embryo-data-plot}", fig.width=6, fig.height=7}
Day <- rep(day,nT)
op <- par(mfrow = c(1,2), lwd = 2)
layout(rbind(c(1, 2), c(3, 3)))
plot(embryo, which = "likelihood")
plot(embryo, which = "change-point")
fk <- density(x = rep((0:20)+.5, nT), bw = "sj", n = 101, from = a, to = b)
hist(Day, breaks = seq(a,b,  length = 12), freq = FALSE, col = "grey", border = "white", 
     main = "Histogram and Density Estimates")
plot(embryo, which = "density", cex = 0.7, add = TRUE)
lines(fk, lty = 2, col = 2)
legend("top", lty = c(1:2), c("MABLE", "Kernel"), bty = "n", col = c(1:2))
par(op)
```
We see in Figure \ref{fig:chicken-embryo-data-AIC-BIC-plot} that AIC and gamma change-point method give the same optimal degree as the one, $m=$ `r embryo$m`, given by the exponential change-point method. However, BIC fails in choosing a useful model degree. 
```{r fig.align='center', fig.cap="AIC and BIC based on Chicken Embryo Data\\label{fig:chicken-embryo-data-AIC-BIC-plot}", fig.width=6, fig.height=4}
M <- embryo$M[1]:embryo$M[2]
aic <- embryo$ic$AIC
bic <- embryo$ic$BIC
res.gcp <- optim.gcp(embryo) # choose m by gamma change-point model
lr <- res.gcp$lr
plot(M, aic, cex = 0.7, col = 1, xlab = "m", ylab = "", ylim = c(ymin<-min(aic,bic,lr),
    ymax<-max(aic,bic,lr)), main = "AIC, BIC, and LR")
points(M, bic, pch = 2, cex = 0.7, col = 2)
points(M[-1], lr, pch = 3, cex = 0.7, col = 4)
segments(which.max(aic)+M[1]-1->m1, ymin, m1, max(aic), lty = 2)
segments(which.max(bic)+M[1]-1->m2, ymin, m2, max(bic), lty = 2, col = 2)
segments(which.max(lr)+M[1]->m3, ymin, m3, max(lr), lty = 2, col = 4)
axis(1, c(m1,m2, m3),  as.character(c(m1,m2,m3)), col.axis = 4)
legend("right", pch = 1:3, c("AIC", "BIC", "LR"), bty = "n", col = c(1,2,4))
```
The results are summarized as follows.
```{r}
summary(embryo)
```

## Contaminated Data--Density Deconvolution

Consider the additive measurement error model $Y = X + \epsilon$, where $X$ has an unknown distribution $F$, $\epsilon$ has a known distribution $G$, and $X$ and $\epsilon$ are independent. We want to estimate density $f=F'$ based on independent observations, $y_i = x_i + \epsilon_i$, $i=1,\ldots,n$, of $Y$, where $x_i$'s and $\epsilon_i$'s are unobservable. \texttt{mable.decon()} implements the method of @Guan-2019-mable-deconvolution and gives an estimate of the density $f$ using the approximate Bernstein polynomial model.

### Example: A Simulated Normal Dataset
```{r results = "hide", warning = FALSE}
set.seed(123)
mu <- 1; sig <- 2; a <- mu - sig*5; b <- mu + sig*5;
gn <- function(x) dnorm(x, 0, 1)
n <- 50;
x <- rnorm(n, mu, sig); e <- rnorm(n); y <- x + e;
decn <- mable.decon(y, gn, interval = c(a,b), M = c(5, 50))
```

```{r fig.align='center', fig.cap="Simulated Normal Data\\label{fig:simulated-normal-data-plot}", fig.width=7, fig.height=7}
op <- par(mfrow = c(2,2), lwd = 2)
plot(decn, which = "likelihood")
plot(decn, which = "change-point", lgd.x = "right")
plot(xx<-seq(a, b, length = 100), yy<-dnorm(xx, mu, sig), type = "l", xlab = "x",
     ylab = "Density", ylim = c(0, max(yy)*1.1))
plot(decn, which = "density", add = TRUE, lty = 2, col = 2)
# kernel density based on pure data
lines(density(x), lty = 5, col = 4)
legend("topright", bty = "n", lty = c(1,2,5), col = c(1,2,4), c(expression(f),     
    expression(hat(f)),  expression(tilde(f)[K])))
plot(xx, yy<-pnorm(xx, mu, sig), type = "l", xlab = "x", ylab = "Distribution Function")
plot(decn, which = "cumulative", add = TRUE, lty = 2, col = 2)
legend("bottomright", bty = "n", lty = c(1,2), col = c(1,2), c(expression(F), 
    expression(hat(F))))
par(op)
```

## Interval Censored Data

When data are interval censored, the ``\texttt{interval2}'' type observations are of the form $(l,u)$ which is the interval containing the event time. Data is uncensored if $l = u$, right censored if $u =$ \texttt{Inf}  or $u =$ \texttt{NA}, and  left censored data if $l =0$.  

Let $f(t)$ and $F(t)=1-S(t)$ be the density and cumulative distribution functions of the event time, respectively, on $(0, \tau)$, where $\tau\le \infty$. If $\tau$ is unknown or $\tau=\infty$, then  $f(t)$ on $[0,\tau_n]$ can be approximated by $fm(t; p)=\tau_n^{-1}\sum_{i=0}^m p_i\beta_{mi}(t/\tau_n)$, where $p_i\ge 0$, $i=0,\ldots,m$, $\sum_{i=0}^mp_i=1-p_{m+1}$, and $\tau_n$ is the largest observation, either uncensored time, or right endpoint of interval/left censored, or left endpoint of right censored time. So we can approximate $S(t)$ on $[0,\tau]$ by $Sm(t; p)=\sum_{i=0}^{m+1} p_i \bar B_{mi}(t/\tau)$, where  $\bar B_{mi}(u)=1-\int_0^u\beta_{mj}(t)dt$, $i=0,\ldots,m$, is the beta survival function with shapes $(i+1,m-i+1)$, $\bar B_{m,m+1}(t)=1$, $p_{m+1}=1-\pi$, and $\pi=F(\tau_n)$. For data without right-censored time, $p_{m+1}=1-\pi=0$. The search for optimal degree $m$ among \texttt{M=c(m0,m1)} using the method of change-point is stopped if either \texttt{m1} is reached or the test for change-point results in a p-value \texttt{pval} smaller than \texttt{sig.level}. @Guan-SIM-2020 proposed a method, as a special case of a semiparametric regression model, for estimating $p$ with an optimal degree $m$. The method is implemented in R function \texttt{mable.ic()}.  

### Example: The Breast Cosmesis Data
Consider the breast cosmesis data as described in @Finkelstein-and-Wolfe-1985-biom is used to study the cosmetic effects of cancer therapy. The time-to-breast-retractions in months ($T$) were  subject to interval censoring and were measured for 94 women among them 46 received radiation only ($X=0$) (25 right-censored, 3 left-censored and 18 interval censored) and 48 received radiation plus chemotherapy ($X=1$)  (13 right-censored, 2 left-censored and 33 interval censored). The right-censored event times were for those women who did not experienced cosmetic deterioration.   

Load package \texttt{coxinterval} to access Breast Cosmesis Data. We fit the two-sample data separately.
```{r results = "hide", warning = FALSE}
library(coxinterval) 
```
```{r}
bcos=cosmesis
head(bcos, 3)
```
```{r results = "hide"}
bc.res0 <- mable.ic(bcos[bcos$treat == "RT",1:2], M = c(1,50), IC = "none")
bc.res1 <- mable.ic(bcos[bcos$treat == "RCT",1:2], M = c(1,50), IC = "none")
```
As the warning message suggested, we check the \texttt{pval}. The \texttt{pval} when the search stopped is `r round(bc.res0$pval[length(bc.res0$pval)],4)`. 
```{r message=FALSE, warning=FALSE, fig.align='center', fig.cap="Breast Cosmesis Data\\label{fig:Breast-Cosmesis-Data-plot}", fig.width=7, fig.height=7, warning = FALSE}
op <- par(mfrow = c(2,2),lwd = 2)
plot(bc.res0, which = "change-point", lgd.x = "right")
plot(bc.res1, which = "change-point", lgd.x = "right")
plot(bc.res0, which = "survival", xlab = "Months", ylim = c(0,1), main = "Radiation Only")
plot(bc.res1, which = "survival", xlab = "Months", main = "Radiation and Chemotherapy")
par(op)
```

## Multivariate Data
A $d$-variate density $f$ on a hyperrectangle $[a,b]=[a_1,b_1]\times \cdots\times[a_d,b_d]$ can be approximated by a mixture of $d$-variate beta densities on $[a,b]$, $\beta_{mj}(x; a, b)=\prod_{i=1}^d\beta_{m_i,j_i}[(x_i-a_i)/(b_i-a_i)]/(b_i-a_i)$, with proportions $p(j_1,\ldots,j_d)$, $0\le j_i\le m_i, i=1,\ldots,d$. Because all the marginal densities can be approximated by Bernstein polynomials, we can choose optimal degree $m_i$ based on observations of the $i$-th component of $x$. For the $i$-th marginal density, an optimal degree is selected using \texttt{mable()}. Then fit the data using EM algorithm with the selected optimal degrees $m=(m_1,\ldots,m_d)$ to obtain a vector $p$ of the mixture proportions $p(j_1,\ldots,j_d)$, arranged in the lexicographical order of $j = (j_1,\ldots,j_d)$, $(p_0,\ldots,p_{K-1})$, where $K=\prod_{i=1}^d (m_i+1)$. The proposed method of @Wang-and-Guan-2019 is implemented by function \texttt{mable.mvar()}.

### Example: The Old Faithful Data

```{r}
data(faithful)
head(faithful, 3)
```
```{r message=FALSE, warning=FALSE, results = "hide"}
a <- c(0, 40); b <- c(7, 110)
#faith2 <- mable.mvar(faithful, M = c(60,30), interval = cbind(a,b))
faith2 <- mable.mvar(faithful, M = c(46,19), search =FALSE, interval = cbind(a,b))
```
```{r fig.align='center', fig.cap="Density Estimate based on the Old Faithful Data\\label{fig:Old-Faithful-Data-plot}", fig.width=6, fig.height=6, warning = FALSE}
plot(faith2, which = "density")
```
The density surface for two-dimensional data can be plot using the \texttt{plot} method (see Figure\ref{fig:Old-Faithful-Data-plot}). The summarized results are given below.
```{r}
summary(faith2)
```

# Event Time Data with Covariates
Let $T$ be an event time and $X$ be an associated $d$-dimensional covariate with distribution $H(x)$ on $\cal{X}$. We denote the marginal and the conditional survival functions of $T$, respectively, by
 $S(t) = \bar F(t)= 1- F(t)=\Pr(T > t)$ and $S(t | x) = \bar F(t | x)=  1- F(t |  x)=\Pr(T > t |  X=x).$
Let $f(t |  x)$ denote the conditional density of a continuous $T$ given $X=x$. The conditional cumulative hazard function and odds ratio,   respectively,
 $\Lambda(t |  x)=-\log S(t |  x)$ and $O(t |  x)={S(t |  x)}/[{1-S(t |  x)}]$.
 We will consider the general situation where the event time is subject to interval censoring.  The observed data are $(X, Y)$, where $Y=(Y_1,Y_2]$, $0\le Y_1\le Y_2\le\infty$.
The reader is referred to @Huang-and-Wellner-1997 for a review and more references about interval censoring.
Special cases are right-censoring $Y_2=\infty$, left-censoring $Y_1=0$, and current status data.

## Cox Proportional Hazards Model

 Consider the Cox proportional hazard regression model [@Cox1972]
\begin{equation}\label{eq: Cox PH model-conditional survival function}
S(t |  x) =S(t | x; \gamma, f_0)=S_0(t)^{\exp(\gamma^T\tilde{x})},
\end{equation}
where $\gamma\in \mathbb{G}  \subset \mathbb{R}^d$, $\tilde{x} =x -x_0$, $x_0$ is any fixed covariate value, $f_0(\cdot)=f(\cdot | x_0)$ is the unknown baseline density and  $S_0(t)=\int_t^\infty f_0(s)ds$. Define $\tau=\inf\{t: F(t |  x_{0})=1\}$.
It is true that $\tau$ is independent of $x_0$, $0<\tau\le\infty$, and $f(t |  x)$ have the same support $[0,\tau]$ for all $x\in \cal{X}$. 
Let $(x_i, y_i)=(x_i, (y_{i1}, y_{j2}])$, $i=1,\dots,n$, be independent observations of $(X, Y)$ and
$\tau_n\ge y_{(n)}=\max\{y_{i1}, y_{j2}: y_{j2}<\infty;\, i,j=1,\dots,n\}$. 
Given any $x_0$, denote $\pi=\pi(x_0)=1-S_0(\tau_n)$. For integer $m\ge 1$ we define $\mathbb{S}_m\equiv \{(u_0,\ldots,u_m)^T\in \mathbb{R}^{m+1}: u_i\ge 0, \sum_{i=0}^m u_i=1.\}.$
@Guan-SIM-2020 propose to approximate $f_0(t)$   on $[0,\tau_n]$  by $f_m(t; p) = \tau_n^{-1}\sum_{i=0}^{m} p_i\beta_{mi}(t/\tau_n)$, where $p=p(x_0)=(p_0,\ldots,p_{m+1})$  are subject to constraints   $p\in \mathbb{S}_{m+1}$ and $p_{m+1}=1-\pi$.
 Here the dependence of $\pi$ and $p$ on $x_0$ will be suppressed.
If $\pi< 1$,  although we cannot estimate the values of $f_0(t)$  on $(\tau_n,\infty)$, we can put an arbitrary guess on them such as
$f_m(t; p)=  p_{m+1} \alpha(t-\tau_n)$, $t\in (\tau_n,\infty)$, where  $\alpha(\cdot)$ is a density on $[0,\infty)$ such that $(1-\pi)\alpha(0)=(m+1)p_m/\tau_n$ so that $f_m(t; p)$ is continuous at $t=\tau_n$, e.g., $\alpha(t)=\alpha(0)\exp[-\alpha(0)t]$.  If $\tau$ is finite and known we choose $\tau_n=\tau$ and specify $p_{m+1}=0$. Otherwise,  we choose $\tau_n=y_{(n)}$. For data without right-censoring or covariate
we also have to specify $p_{m+1}=0$ due to its unidentifiability. 

The above method is implemented in function \texttt{mable.ph()} which returns maximum approximate Bernstein likelihood estimates of $(\gamma, p)$ with an optimal model degree $m$ and a prespecified $m$,  respectively. With a efficient estimate of $\gamma$ obtained using other method, \texttt{maple.ph()} can be used to get an optimal degree $m$ and a mable of $p$. 

The \texttt{plot} method \texttt{plot.mable\_reg()} for class \texttt{mable\_reg} object returned by all the above functions produces graphs of the loglikelihoods at $m$ in a set \texttt{M[1]:M[2]} of consecutive candidate model degrees, the likelihood ratios of change-point at $m$ in \texttt{(M[1]+1):M[2]},  estimated density and survival function on the truncated \texttt{support}=$[0,\tau_n]$.

### Example: Ovarian Cancer Survival Data
The Ovarian Cancer Survival Data is included in package \texttt{survival}.
```{r}
futime2 <- ovarian$futime
futime2[ovarian$fustat==0] <- Inf
ovarian2 <- data.frame(age = ovarian$age, futime1 = ovarian$futime, futime2 = futime2)
head(ovarian2, 3)
```
```{r results = "hide", warning = FALSE, message=FALSE}
ova<-mable.ph(cbind(futime1, futime2) ~ age, data = ovarian2, M = c(2,35), g = .16, x0=35)
```
```{r}
summary(ova)
```
```{r fig.align='center', fig.cap="Ovarian Cancer Data\\label{fig:ovarian-Data-plot}", fig.width=7, fig.height=7, warning = FALSE, message=FALSE}
op <- par(mfrow = c(2,2))
plot(ova, which = "likelihood")
plot(ova, which = "change-point")
plot(ova, y=data.frame(age=60), which="survival", type="l", xlab="Days", main="Age = 60")
plot(ova, y=data.frame(age=65), which="survival", type="l", xlab="Days", main="Age = 65")
par(op)
```
Alternatively we can use \texttt{mable.reg()}.
```{r results = "hide", warning = FALSE}
ova1 <- mable.reg(cbind(futime1, futime2) ~ age, data = ovarian2, M = c(2,35))
```

## Accelerated Failure Time Model

The AFT model can be specified as
\begin{equation}\label{eq: AFT model in terms of density}
f(t\mid x)=f(t\mid x;\gamma)=e^{-\gamma^T x}f(te^{-\gamma^T x}\mid 0),\quad t\in[0,\infty),
\end{equation}
where $\gamma\in \mathbb{G}\subset \mathbb{R}^d$. Let $\gamma_0\in \mathbb{G}$ be the true value of $\gamma$.
The AFT model (\ref{eq: AFT model in terms of density}) is equivalent to
$$S(t\mid x;\gamma)=S(te^{-\gamma^T x}\mid 0),\quad t\in[0,\infty).$$
Thus this is actually a scale regression model.
The AFT model can also be written as linear regression $\log(T)=\gamma^T x +\varepsilon$. It is clear that one can choose any $x_0$ in $\mathcal{X}$ as baseline by transform $\tilde{x}=x-x_0$. If $f(t\mid 0)$ has support $[0,\tau_0)$, $\tau_0\le\infty$, then 
$f(t\mid x)$ has support $[0,\tau_0 e^{\gamma_0^T x})$.  We define $\tau=\max\{\tau_0 e^{\gamma_0^T x}: x\in\mathcal{X}\}$ if $\tau_0<\infty$ and $\tau=\infty$ otherwise.
The above AFT model can also be written as
$$f(t\mid x;\gamma)=e^{-\gamma^T {x}} f_0(te^{-\gamma^T {x}}), \quad
S(t\mid x;\gamma)=S_0(te^{-\gamma^T {x}}),$$
where   $f_0(t)=f(t\mid 0)$ and $S_0(t)=S(t\mid 0)=\int_t^\infty f_0(u)du$.

As in Cox PH model, we choose $\tau_n>y_{(n)}= \max\{y_{i1}, y_{j2}: y_{j2}<\infty;\, i,j=1,\dots,n\}$ so that $S(\tau_n)$ and $\max_{x\in\mathcal{X}} S(\tau_n\mid x)$ are believed very small.
Then we approximate $f_0(t)$ and $S_0(t)$ on $[0,\tau_n]$, respectively, by
\begin{align*}
f_0(t)&\approx f_m(t; p)=\frac{1}{\tau_n}\sum_{j=0}^m p_j\beta_{mj}\Big(\frac{t}{\tau_n}\Big),\quad t\in[0,\tau_n];\\
S_0(t)&\approx S_m(t; p)= \sum_{j=0}^{m} p_j \bar B_{mj}\Big(\frac{t}{\tau_n}\Big),\quad  t\in[0,\tau_n],
\end{align*}
where $S_m(\infty; p)=0$, and $p= (p_0,\ldots,p_{m})^T\in \mathbb{S}_{m}$.
  Then $f(t\mid x; \gamma)$ and $S(t\mid x; \gamma)$  can be approximated, respectively, by
\begin{align}\nonumber
f_m(t\mid x; \gamma, p)&=e^{-\gamma^T {x}}f_m\Big(te^{-\gamma^T {x}}; p\Big)\\
&=
 \frac{e^{-\gamma^T {x}}}{\tau_n}\sum_{j=0}^m p_j\beta_{mj}\Big(e^{-\gamma^T {x}}\frac{t}{\tau_n}\Big),\quad t\in[0,\tau_ne^{\gamma^T {x}}];\\\nonumber
S_m(t\mid x;\gamma,p)&=S_m\Big(te^{-\gamma^T {x}}; p\Big)\\
&=  \sum_{j=0}^{m} p_j\bar B_{mj}\Big(e^{-\gamma^T {  x}}\frac{t}{\tau_n}\Big),\quad t\in[0,\tau_ne^{\gamma^T {x}}].
\end{align}
@Guan-2019-mable-aft's proposal is implemented by functions \texttt{mable.aft()} and \texttt{maple.aft()}.

### Example: Breast Cosmesis Data

```{r}
bcos2 <- data.frame(bcos[,1:2], x = 1*(bcos$treat == "RCT"))
```
```{r results = "hide", warning=FALSE, message=FALSE}
g <- 0.41 # Hanson and Johnson 2004, JCGS
aft.res<-mable.aft(cbind(left, right) ~ x, data = bcos2, M =c(1, 30), g=.41,  tau =100, x0=1)
```
```{r}
summary(aft.res)
```
```{r fig.align='center', fig.cap="AFT Model Fit for Breast Cosmesis Data\\label{fig:Breast-Cosmesis-Data-aft-plot}", fig.width=7, fig.height=4, warning = FALSE}
op <- par(mfrow = c(1,2), lwd = 1.5)
plot(x = aft.res, which = "likelihood")
plot(x = aft.res, y = data.frame(x = 0), which = "survival", type = "l", col = 1, 
    main = "Survival Function")
plot(x = aft.res, y = data.frame(x = 1), which = "survival", lty = 2, col = 1, add = TRUE)
legend("bottomleft", bty = "n", lty = 1:2, col = 1, c("Radiation Only", 
    "Radiation and Chemotherapy"), cex = .7)
par(op)
```
Alternatively we can use \texttt{mable.reg()}.
```{r results = "hide"}
aft.res1 <- mable.reg(cbind(left, right) ~ x, data = bcos2, 'aft', M = c(1, 30), 
      g=.41, tau=100, x0=1)
```

# Two-sample Data
In addition to the PH and AFT regression models with single binary covariate, other two-sample semiparametric models can be estimated using the maximum approximate Bernstein/Beta likelihhod method.

## Density Ratio Model

Two-sample density ratio(DR) model
(see for example, @QinZhang1997 and @Qin-and-Zhang-2005) is useful for fitting case-control data.
 Suppose that the densities $f_0$ and $f_1$ of $X_0$ and $X_1$, respectively, satisfy the following density ratio model (see @CHENG-and-CHU-Bernoulli-2004 and @Qin-and-Zhang-2005, for example)
\begin{equation}\label{eq: exponential tilting model}
f_1(x)=f(x;\bm\alpha)=f_0(x)\exp\{\bm\alpha^T \tilde r(x)\},
\end{equation}
where $\bm\alpha=(\alpha_0,\ldots,\alpha_d)^T \in\mathcal{A}\subset R^{d+1}$, and $\tilde r(x)=(1, r^T(x))^T$ with linearly independent components. @Guan-arXiv-2021-mable-dr-model proposed to approximate the \emph{baseline} density $f_0$ by Bernstein polynomial and to estimate $f_0$ and $\bm\alpha$ using the maximum approximate Bernstein/Beta likelihood estimation. This method is implemented by function \texttt{mable.dr()} and \texttt{maple.dr()} in which $\bm\alpha$ is a given estimated value such as the one obtained by the logistic regression as described in @QinZhang1997 and @Qin-and-Zhang-2005.

### Example: Coronary Heart Disease Data 
```{r message=FALSE, warning = FALSE, results = "hide"}
# Hosmer and Lemeshow (1989):                                           
# ages and the status of coronary disease (CHD) of 100 subjects         
x<-c(20, 23, 24, 25, 26, 26, 28, 28, 29, 30, 30, 30, 30, 30, 32,        
32, 33, 33, 34, 34, 34, 34, 35, 35, 36, 36, 37, 37, 38, 38, 39,         
40, 41, 41, 42, 42, 42, 43, 43, 44, 44, 45, 46, 47, 47, 48, 49,         
49, 50, 51, 52, 55, 57, 57, 58, 60, 64)                                 
y<-c(25, 30, 34, 36, 37, 39, 40, 42, 43, 44, 44, 45, 46, 47, 48,        
48, 49, 50, 52, 53, 53, 54, 55, 55, 56, 56, 56, 57, 57, 57, 57,         
58, 58, 59, 59, 60, 61, 62, 62, 63, 64, 65, 69) 
a<-20; b<-70
regr<-function(x) cbind(1,x)                                            
chd.mable<-mable.dr(x, y, M=c(1, 15), regr, interval = c(a,b))   
```
```{r fig.align='center', fig.cap="DR Model Fit for Coronary Heart Disease  Data\\label{fig:CHD-data-plot}", fig.width=7, fig.height=4, warning = FALSE}
z<-seq(a,b,length=512)
f0hat<-dmixbeta(z, p=chd.mable$p, interval=c(a, b))
rf<-function(x) chd.mable$regr((x-a)/(b-a))
f1hat<-dtmixbeta(z, p=chd.mable$p, alpha=chd.mable$alpha, 
                 interval=c(a, b), regr=rf)
op<-par(mfrow=c(1,2),lwd=1.2, cex=.7, mar=c(5,4,1,1))
hist(x, freq=F, col = "light grey", border = "white", xlab="Age", 
  ylab="Density", xlim=c(a,b), ylim=c(0,.055), main="Control")
lines(z, f0hat, lty=1, col=1)
hist(y, freq=F, col = "light grey", border = "white", xlab="Age", 
  ylab="Density", xlim=c(a,b), ylim=c(0,.055), main="Case")
lines(z, f1hat, lty=1, col=1)
par(op)
```


### Example: Pancreatic Cancer Biomarker Data
For the logarithmic levels of CA 19-9 of the Pancreatic Cancer Data (@Wieand-et-al-1989-bka), the control group is used as baseline because the optimal model degree is smaller that the one using case as baseline.  
```{r message=FALSE, warning = FALSE, results = "hide"}
data(pancreas)
head(pancreas,3)
x<-log(pancreas$ca199[pancreas$status==0])
y<-log(pancreas$ca199[pancreas$status==1])
a<-min(x,y); b<-max(x,y)
M<-c(1,29)
regr<-function(x) cbind(1,x,x^2)                                            
m=maple.dr(x, y, M, regr=regr, interval=c(a,b), controls=mable.ctrl(sig.level=.001))$m
pc.mable<-mable.dr(x, y, M=m, regr=regr, interval=c(a,b),
                  controls=mable.ctrl(sig.level=1/length(c(x,y))))
#pc.mable   
```
```{r fig.align='center', fig.cap="DR Model Fit for Pancreatic Cancer Biomarker Data\\label{fig:pcb-data-plot}", fig.width=7, fig.height=3, warning = FALSE}
z<-seq(a,b,length=512)
# baseline is "case"
f1hat<-dmixbeta(z, p=pc.mable$p, interval=c(a, b))
rf<-function(x) pc.mable$regr((x-a)/(b-a))
f0hat<-dtmixbeta(z, p=pc.mable$p, alpha=pc.mable$alpha, 
                 interval=c(a, b), regr=rf)
op<-par(mfrow=c(1,2),lwd=1.2, cex=.7, mar=c(5,4,1,1))
hist(x, freq=F, col = "light grey", border = "white", xlab="log(CA19-9)", 
  ylab="Density", xlim=c(a,b),  main="Control")
lines(z, f0hat, lty=1, col=1)
hist(y, freq=F, col = "light grey", border = "white", xlab="log(CA19-9)", 
  ylab="Density", xlim=c(a,b), ylim=c(0,.5), main="Case")
lines(z, f1hat, lty=1, col=1)
par(op)
```




## References